Learning by Teaching a Robot: the Case of Handwriting

### 2015-07-09 main.tex
School studies section

@@

\paragraph{Engagement levels}

A conclusion drawn in a systematic review of classical handwriting
interventions~\cite{Hoy2011} is that any of the studies considered which
involved fewer than two practice sessions per week and fewer than a total of 20
practice sessions, including homework, were not found to demonstrate effective
results. This highlights the necessity to engage students in an interaction
which will be sustainable over the long-term if we want to address research
questions which involve the measurement of learning gains.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{engagement}
    \caption{Engagement level of the child, manually annotated by two judges
        over the 13 min long interaction (Cronbach’s alpha test: $\alpha = .82$,
        good inter-judge agreement). The vertical lines represent activity
    switches between handwriting teaching and story-telling (used as a
    distruber).}
    \label{}
\end{figure}

>>>>

Over the two years of the project, we conducted four studies in schools
(Figure~\ref{fig:schools}). These experiments were meant to technically validate
the system (is it actually able to autonomously write and learn from
demonstrations?) and the interaction (is the apparatus easy to grasp and to
interact with for children), and to study the initial acceptance of the robot in
the school environment (we conducted several formal and informal discussions
with teachers).

We also relied on these four studies to acquire a dataset of children's
handwritten letters and to study how they engage with the robot (and maintain or
not this engagement).

@@

Critically, these studies were conducted with \emph{whole classes}: we did not
select underperforming children that requires careful beforehand organization
with the school to limit ethical issues. Having more children (73 in total) was
however beneficial for these preliminary studies.

\paragraph{System validation}

From a technical perspective, the system achieves an acceptable level of
reliability and allow a technically sound autonomous interaction. For instance,
during the second school study (\textit{Florimont}), the robot withstood
interactions which lasted for a total of 160 minutes.  During this time the
robot wrote 335 letters, 152 of which in response to demonstrations received
from the 21 children. Technical intervention was only required for the three
instances that the robot fell later in that day.  Otherwise, the technical
components of the system operated autonomously and as expected over the
sessions.

Due to the modular software architecture (about 10 ROS nodes), crashes occurring
during the sessions were usually quickly resolved by re-launching the faulty
node alone, and did not significantly impact the interaction.

The otherwise technical limitations were related to some letters or writing
style (most notably, the ones requiring multiple strokes per letter) not being
adequately processed by the learning algorithm. Support for such letters has
been since added.

\paragraph{Acceptance}

For the approach to be effective, it is critical that children recognize and
accept that the robot is writing by itself. When asked, no child indicated that
they did not believe that the robot was writing by itself. There were, at times,
questions about the robot's writing method at the beginning of the interaction,
but when advised that the robot ``tells the tablet what it wants to write,''
this was accepted by the children.  Besides, teachers interviewed for their
feedback on the system advised that children are asked to draw letters in the
air in a similar manner as part of their handwriting education. The behaviour is
hence not unfamiliar to children.

\paragraph{Engagement}

As indicated in the introduction, literature~\cite{Hoy2011} suggests that a
total of 20 handwriting practice sessions is found to be the minimum to demonstrate effective
results in handwriting remediation. This highlights the necessity to sustain
engagement over the long-term if we want to achieve measurable learning gains.

Factually, the children engaged into the teaching activity: in the
\textit{Florimont} study, for instance, they demonstrated an average of
10.9 demonstration letters (SD = 4.4) for an average session duration of 11 minutes.
In 9 out of the 14 sessions (64\%), the robot
received demonstration letters even \emph{after} reaching the final stage of the
interaction, suggesting an intrinsic motivation to further engage in the
interaction.

\begin{figure}[ht!]
    \centering
    \includegraphics[width=0.9\linewidth]{engagement}
    \caption{Engagement level of one child, manually annotated by two judges
        over a 13 min long interaction (Cronbach’s alpha test: $\alpha = .82$,
        good inter-judge agreement). 0 means ``complete disengagement from the
        task'', 6 means ``full engagement''. The vertical lines represent activity
    switches between handwriting teaching and story-telling (used as a
    distruber).}
    \label{engagement_level}
\end{figure}

We also conducted a qualitative assessment of the engagement levels during the
interaction sessions (studies \textit{Haut-Lac} and \textit{ISG 2}).
Figure~\ref{engagement_level} shows the result for one child, with a moderate
level of engagement. On-going research in our group aims at assessing such
engagement levels in real-time, and adapt accordingly its behaviour.

Other qualitative assessment of engagement over longer interactions were
conducted during the case studies, as reported in the following section.
